---
title: "Learn Time Series Week 1"
output: html_notebook
---
I'm working through the material in the coursera time series analysis course.

Create some data
```{r}
data<-c(35,8,10,23,42)
```

Five number summary
```{r}
summary(data)
mean(data)
sd(data)
```

Histogram
```{r}
small.size.dataset<-c(91,49,76,112,97,47, 70, 100, 8, 112, 95, 90, 78, 62, 56, 94, 65, 58, 109, 70, 109, 91, 71, 76, 68, 62, 134, 57, 83, 66)
```
create histogram using base r
```{r}
hist(small.size.dataset)
```
Make the plot nicer.
* change x-axis label
```{r}
hist(small.size.dataset, xlab = "My Data Points")
```
* Change histogram title
```{r}
hist(small.size.dataset, xlab = "My Data Points", main = "Histogram of my data")
```

* Change y-axis to probability or density.
```{r}
hist(small.size.dataset, xlab = "My Data Points", freq = F)
```
* change colors
```{r}
hist(small.size.dataset, xlab = "My Data Points", col = "green")
```
Now put a smooth density function over the histogram
```{r}
hist(small.size.dataset, xlab = "My Data Points", col = "green")
lines(density(small.size.dataset))
```
Now some basic review of linear regression
```{r}
co2 #data from Mauna Loa atmostpheric CO2 concentration, built in with R
help(co2)
#what kind of object is the data?
class(co2)
```
Make a plot
```{r}
plot(co2, main = "Atmospheric CO2 concentration")
```
Calculate slope and intercept by hand
```{r}
co2.values<-as.numeric(co2)
co2.times = as.numeric(time(co2))
```
Now calculate sums of squares
```{r}
SSxx<-sum((co2.times - mean(co2.times))*(co2.times -mean(co2.times)))
SSxy<-sum((co2.values - mean(co2.values))*((co2.times-mean(co2.times))))
```
And now the slope and intercept
```{r}
slope<-SSxy/SSxx
intercept<-mean(co2.values)-slope*mean(co2.times)
```
So we just made a linear model by hand.  R can do it quickly for us:
```{r}
co2.linear.model<-lm(co2~time(co2))
```
And now place the fitted line ona a graph
```{r}
plot(co2, main = "Atmospheric CO2 concentration with fitted line")
abline(co2.linear.model)
```
And calculate the residuals by hand:
```{r}
co2.fitted.values<-slope*co2.times + intercept
co2.residuals<-co2.values - co2.fitted.values
```
Or we can let R do it for us:
```{r}
co2.residualsR<-resid(co2.linear.model)
```
An assumption of linear regression is that the residuals are normally distributed.  Let's check, by plotting a histogram (with a large data set) or a normal probability plot (with a smaller data set). Departures from a straight line in a probability plot indicate likely departures from normality.

It is also common to plot residuals against the time (predicted) variable.

```{r}
par(mfrow = c(1, 3))
hist(co2.residualsR, main = "Histogram of CO2 residuals")
qqnorm(co2.residualsR, main = "Normal probability plot")
qqline(co2.residualsR)
plot(co2.residualsR ~ time(co2), main = "Residuals on time")
```


